### 2021-10-26 publication/experiments.tex
cosmetic changes

	\item \textbf{Primary school dynamic contacts} \cite{schools} ($N=238, E=5539, D=13$) -- network of face-to-face contacts amongst students and teachers at a primary school in Lyon, France. Two nodes are connected if the two parties shared a face-to-face interaction over the school-day. Vertex features include class membership (one of 10 values: 1A-5B), gender (male, female) and teacher status encoded as an 11th school-class. No further identifiable information is retained. We choose to analyse just the second day of results.

@@

\begin{figure}[!h]
	\centering
	\begin{subfigure}[t]{0.28\linewidth}
		\centering
		\includegraphics[width=\linewidth]{polbooks-graph.png}
		\caption{Polbooks}
		\label{fig:polbooks-graph}
	\end{subfigure}
	\hfill
	\begin{subfigure}[t]{0.28\linewidth}
		\centering
		\includegraphics[width=\linewidth]{school-graph.png}
		\caption{School}
		\label{fig:school-graph}
	\end{subfigure}
	\hfill
	\begin{subfigure}[t]{0.28\linewidth}
		\centering
		\includegraphics[width=\linewidth]{fb-graph.png}
		\caption{Facebook egonet}
		\label{fig:fb-graph}
	\end{subfigure}
	\begin{subfigure}[t]{0.11\linewidth}
		\centering
		\includegraphics[width=0.8\linewidth]{10-vertical-legend.png}
		\caption{Legend}
		\label{fig:10-legend}
	\end{subfigure}
	\caption{Networks laid out and coloured according to inferred block memberships $\hat{y}$ for a given experiment iteration. Visualisation performed using \textit{graph-tool} \cite{peixoto_graph-tool_2014}.}
	\label{fig:graphs-all}
\end{figure}
\begin{table}[!h]
	\centering
	\caption{Experimental results averaged over $n=10$ iterations (mean $\pm$ standard deviation).}
	\label{tab:results}
	\resizebox{\textwidth}{!}{%
		\begin{tabular}{c|ccc|c|cc|ccc}
			Dataset  & $B$ & $D$ & $D'$ & $\bar{S}_e$ & $\bar{\Lcal}_0$ & $\bar{\Lcal}_1$ & $c^*$ & $\bar{\Lcal}_0'$ & $\bar{\Lcal}_1'$  \\ \hline
			Polbooks & 3 & 3 & -- & $2.250 \pm 0.000$ & $0.563 \pm 0.042$ & $0.595 \pm 0.089$ & -- & -- & -- \\
			School & 10 & 13 & 10 & $1.894 \pm 0.004$ & $0.787 \pm 0.127$ & $0.885 \pm 0.129$ & $1.198 \pm 0.249$ & $0.793 \pm 0.132$ & $0.853 \pm 0.132$ \\
			FB egonet & 10  & 480 & 10 & $1.626 \pm 0.003$ & $1.326 \pm 0.043$ & $1.538 \pm 0.069$ & $0.94 \pm 0.019$ & $1.580 \pm 0.150$ & $1.605 \pm 0.106$
		\end{tabular}
	}
\end{table}
%
We require metrics to assess performance. This can be split into two separate 
components: the microcanonical SBM fit (concerned with the $b$-samples) and 
the fit of the feature-to-block generator (concerned with the $\theta$-samples).

@@

Table~\ref{tab:results} summarises the results for each experiment;
a complete list of all the hyperparameters used 
is given in Appendix \ref{appdx:hyperparams}. We also apply the 
dimensionality reduction method 
of Section~\ref{sec:dim-reduction}
to the two higher dimensional datasets (the school and FB egonet). 
For this we use equation~(\ref{eqn:c-star}) with $k=1$,
in order to reduce the dimension from 
$D$ to a desired $D'$. 
We then retrain the feature-block predictor using only the retained 
feature set $\Dcal'$, and report the log-loss over the training and 
test sets for the reduced classifier -- 
denoted $\bar{\Lcal}_0'$ and $\bar{\Lcal}_1'$ respectively. 
These values are also given in Table~\ref{tab:results}.

@@

We choose the number of communities $B=10$, in line with the total number of 
school classes. As before, we sample the block-generator parameters $\theta$ and 
employ the dimensionality reduction technique 
of Section~\ref{sec:dim-reduction},
with standard deviation multiplier $k=1$ to pick out the top $D'=10$ features. 
We then plot the weights for the resulting features $d \in \Dcal'$ 
in Figure~\ref{fig:school-null}. It is rather obvious
that only the pupils' class memberships (1A-5B) have remained
significant;
gender and teacher/student status have been discarded,
meaning that these are not good predictors of overall macro-structure.
%
\begin{figure}[!h]
	\centering
	\begin{subfigure}[t]{0.45\linewidth}
		\centering
		\imagebox{6.2cm}{\includegraphics[width=\linewidth]{school-null-1}}
		\caption{$\theta$-samples. Dotted line is $\pm c^*$.}
		\label{fig:school-null}
	\end{subfigure}
	\begin{subfigure}[t]{0.45\linewidth}
		\centering
		\imagebox{6.2cm}{\includegraphics[width=\linewidth]{school-accuracy-1}}
		\caption{Per-block accuracy $\eta(j)$.}
		\label{fig:school-accuracy}
	\end{subfigure}
	\caption{Primary school dynamic contacts dataset.}
	\label{fig:school}
\end{figure}

@@

We choose $B=10$ and $D'=10$ for this experiment. The selected features 
(Figure~\ref{fig:fb-null}) are those that best explain the high-level 
community structure. The majority of them are education related. 
Nevertheless, for $D'=10$ we only have good explanations for the makeup 
of some of the detected blocks; several blocks in 
Figure~\ref{fig:fb-null} do not have high-magnitude components for $D'=10$. This is further emphasised by the disparate accuracies in Figure~\ref{fig:fb-accuracy}. Nevertheless, observe that the accuracy is only high for blocks that contain high magnitude weights. The only exception to this rule is block 9 which nonetheless may have high magnitude weights just below the cut-off $c^*$.
%
\begin{figure}[!h]
	\centering
	\begin{subfigure}[t]{0.45\linewidth}
		\centering
		\imagebox{7cm}{\includegraphics[width=\linewidth]{fb-null-1}}
		\caption{$\theta$-samples. Dotted line is $\pm c^*$.}
		\label{fig:fb-null}
	\end{subfigure}
	\begin{subfigure}[t]{0.45\linewidth}
		\centering
		\imagebox{7cm}{\includegraphics[width=\linewidth]{fb-accuracy-1}}
		\caption{Per-block accuracy $\eta(j)$.}
		\label{fig:fb-accuracy}
	\end{subfigure}
	\caption{Facebook egonet dataset.}
	\label{fig:fb}
\end{figure}

When the feature dimension is very large, it becomes increasingly likely that 
a particular feature may uniquely identify a small set of nodes. If these nodes 
are all part of the same community, then the classifier may overfit for that 
particular parameter. The regularisation term imposed by the prior goes some 
way towards alleviating this problem. Nevertheless, we see in 
Figure~\ref{fig:fb-null} that the feature \verb*|birthday-5| has a very 
high weight as it relates to block 1 -- but it is highly unlikely
that birthdays determine graphical structure. This issue could have been alleviated by 
choosing $\Xcal=\{-1, +1\}$ such that having a feature turned off supports excluding 
a vertex from the block. However, Figure~\ref{fig:fb-null} demonstrates why we made 
the deliberate choice $\Xcal=\{0,1\}$, as argued in Appendix \ref{appdx:dimension}. 
Because of the choice $\Xcal=\{0,1\}$, block 1 is allowed to contain 3 distinct feature 
categories rather than just one. We accept this trade-off: risking spurious high magnitude 
components (e.g. \verb*|birthday-5|) for the benefit of allowing a top-level block to span 
multiple feature-groups. Nevertheless, one would have to apply a hierarchical approach 
to subdivide each top-level block into sub-blocks to test whether the constituent 
feature-groups determine structure at the lower-level.

>>>>

	\item \textbf{Primary school dynamic contacts} \cite{schools} ($N=238, E=5539, D=13$) -- network of face-to-face contacts amongst students and teachers at a primary school in Lyon, France. Two nodes are connected if the two parties shared a face-to-face interaction over the school-day. Vertex features include class membership (one of 10 values: 1A-5B), gender (male, female) and teacher status encoded as an 11th school-class. We choose to analyse just the second day of results.

@@

\begin{figure}[!h]
	\centering
	\begin{subfigure}[t]{0.28\linewidth}
		\centering
		\includegraphics[width=\linewidth]{polbooks-graph.png}
		\caption{Polbooks}
		\label{fig:polbooks-graph}
	\end{subfigure}
	\hfill
	\begin{subfigure}[t]{0.28\linewidth}
		\centering
		\includegraphics[width=\linewidth]{school-graph.png}
		\caption{School}
		\label{fig:school-graph}
	\end{subfigure}
	\hfill
	\begin{subfigure}[t]{0.28\linewidth}
		\centering
		\includegraphics[width=\linewidth]{fb-graph.png}
		\caption{Facebook egonet}
		\label{fig:fb-graph}
	\end{subfigure}
	\begin{subfigure}[t]{0.11\linewidth}
		\centering
		\includegraphics[width=0.8\linewidth]{10-vertical-legend.png}
		\caption{Block Legend}
		\label{fig:10-legend}
	\end{subfigure}
	\caption{Networks laid out and coloured according to inferred block memberships $\hat{y}$ for a given experiment iteration. Visualisation performed using \textit{graph-tool} \cite{peixoto_graph-tool_2014}.}
	\label{fig:graphs-all}
\end{figure}
\begin{table}[!h]
	\centering
	\caption{Experimental results averaged over $n=10$ iterations (mean $\pm$ std. dev.).}
	\label{tab:results}
	\resizebox{\textwidth}{!}{%
		\begin{tabular}{c|ccc|c|cc|ccc}
			Dataset  & $B$ & $D$ & $D'$ & $\bar{S}_e$ & $\bar{\Lcal}_0$ & $\bar{\Lcal}_1$ & $c^*$ & $\bar{\Lcal}_0'$ & $\bar{\Lcal}_1'$  \\ \hline
			Polbooks & 3 & 3 & -- & $2.250 \pm 0.000$ & $0.563 \pm 0.042$ & $0.595 \pm 0.089$ & -- & -- & -- \\
			School & 10 & 13 & 10 & $1.894 \pm 0.004$ & $0.787 \pm 0.127$ & $0.885 \pm 0.129$ & $1.198 \pm 0.249$ & $0.793 \pm 0.132$ & $0.853 \pm 0.132$ \\
			FB egonet & 10  & 480 & 10 & $1.626 \pm 0.003$ & $1.326 \pm 0.043$ & $1.538 \pm 0.069$ & $0.94 \pm 0.019$ & $1.580 \pm 0.150$ & $1.605 \pm 0.106$
		\end{tabular}
	}
\end{table}
%
We require metrics to assess performance. This can be split into two separate 
components: the microcanonical SBM fit (concerned with the $b$-samples) and 
the fit of the feature-to-block generator (concerned with the $\theta$-samples).

@@

Table~\ref{tab:results} summarises the results for each experiment. We also apply the 
dimensionality reduction method 
of Section~\ref{sec:dim-reduction}
to the two higher dimensional datasets (the school and FB egonet). 
For this we use equation~(\ref{eqn:c-star}) with $k=1$,
in order to reduce the dimension from 
$D$ to a desired $D'$. 
We then retrain the feature-block predictor using only the retained 
feature set $\Dcal'$, and report the log-loss over the training and 
test sets for the reduced classifier -- 
denoted $\bar{\Lcal}_0'$ and $\bar{\Lcal}_1'$ respectively. 
These values are also given in Table~\ref{tab:results}.

@@

We choose the number of communities $B=10$, in line with the total number of 
school classes. As before, we sample the block-generator parameters $\theta$ and 
employ the dimensionality reduction technique 
of Section~\ref{sec:dim-reduction},
with standard deviation multiplier $k=1$ to pick out the top $D'=10$ features. 
We then plot the weights for the resulting features $d \in \Dcal'$ 
in Figure~\ref{fig:school-null}. It is rather obvious
that only the pupils' class memberships (1A-5B) have remained
significant;
gender and teacher/student status have been discarded,
meaning that these are not good predictors of overall macro-structure.
%
\begin{figure}[!h]
	\centering
	\begin{subfigure}[t]{0.45\linewidth}
		\centering
		\imagebox{0.8\linewidth}{\includegraphics[width=\linewidth]{school-null-1}}
		\caption{$\theta$-samples. Dotted line is $\pm c^*$.}
		\label{fig:school-null}
	\end{subfigure}
	\begin{subfigure}[t]{0.45\linewidth}
		\centering
		\imagebox{0.8\linewidth}{\includegraphics[width=\linewidth]{school-accuracy-1}}
		\caption{Per-block accuracy $\eta(j)$.}
		\label{fig:school-accuracy}
	\end{subfigure}
	\caption{Primary school dynamic contacts dataset.}
	\label{fig:school}
\end{figure}

@@

We choose $B=10$ and $D'=10$ for this experiment. The selected features 
(Figure~\ref{fig:fb-null}) are those that best explain the high-level 
community structure. The majority of them are education related. 
Nevertheless, for $D'=10$ we only have good explanations for the makeup 
of some of the detected blocks; several blocks in 
Figure~\ref{fig:fb-null} do not have high-magnitude components for $D'=10$. This is further emphasised by the disparate accuracies in Figure~\ref{fig:fb-accuracy}. Nevertheless, observe that the accuracy is only high for blocks that contain high magnitude weights. The only exception to this rule is block 9 which nonetheless may have high magnitude weights just below the cut-off $c^*$.
%
\begin{figure}[!h]
	\centering
	\begin{subfigure}[t]{0.45\linewidth}
		\centering
		\imagebox{0.9\linewidth}{\includegraphics[width=\linewidth]{fb-null-1}}
		\caption{$\theta$-samples. Dotted line is $\pm c^*$.}
		\label{fig:fb-null}
	\end{subfigure}
	\begin{subfigure}[t]{0.45\linewidth}
		\centering
		\imagebox{0.9\linewidth}{\includegraphics[width=\linewidth]{fb-accuracy-1}}
		\caption{Per-block accuracy $\eta(j)$.}
		\label{fig:fb-accuracy}
	\end{subfigure}
	\caption{Facebook egonet dataset.}
	\label{fig:fb}
\end{figure}

When the feature dimension is very large, it becomes increasingly likely that 
a particular feature may uniquely identify a small set of nodes. If these nodes 
are all part of the same community, then the classifier may overfit for that 
particular parameter. The regularisation term imposed by the prior goes some 
way towards alleviating this problem. Nevertheless, we see in 
Figure~\ref{fig:fb-null} that the feature \verb*|birthday-5| has a very 
high weight as it relates to block 1 -- but it is highly unlikely
that birthdays determine graphical structure. This issue could have been alleviated by 
choosing $\Xcal=\{-1, +1\}$ such that having a feature turned off supports excluding 
a vertex from the block. However, Figure~\ref{fig:fb-null} demonstrates why we made 
the deliberate choice $\Xcal=\{0,1\}$. 
Because of the choice $\Xcal=\{0,1\}$, block 1 is allowed to contain 3 distinct feature 
categories rather than just one. We accept this trade-off: risking spurious high magnitude 
components (e.g. \verb*|birthday-5|) for the benefit of allowing a top-level block to span 
multiple feature-groups. Nevertheless, one would have to apply a hierarchical approach 
to subdivide each top-level block into sub-blocks to test whether the constituent 
feature-groups determine structure at the lower-level.

### 2021-10-26 publication/inference.tex
cosmetic changes

Having completed the definition of the FFBM, we wish to leverage it 
to perform inference. Given a vertex-labelled graph $(A, X)$,
the goal is to sample from the posterior distribution
of $\theta$:
%
\begin{equation}
	\label{eqn:theta-target}
	\theta \sim p(\theta | A, X).
\end{equation}
%
We propose an iterative approach to obtain samples
$\theta^{(t)}$ from $p(\theta|A,X)$. We first draw samples $b^{(t)}$ 
from the block membership posterior distribution,
and then use each $b^{(t)}$ to obtain a corresponding
sample $\theta^{(t)}$:
%
\begin{align}
	b^{(t)} &\sim p ( b | A, X )  \label{eqn:b-samples},\\
	\theta^{(t)} &\sim p(\theta | X, b^{(t)} ). \label{eqn:theta-samples}
\end{align}
%
Both of these sampling steps can be implemented through a
Markov chain Monte-Carlo approach via a Metropolis-Hastings 
algorithm \cite{hastings-alg}, as a two-level Markov chain. See Appendix \ref{appdx:metropolis-hastings} for more details on the Metropolis-Hastings algorithm.

@@

	\begin{tikzpicture}[
		scale=0.95, every node/.style={transform shape},
		roundnode/.style={circle, draw=black, minimum size=12mm},
		squarednode/.style={rectangle, draw=black, minimum size=12mm}
		]
		% nodes
		\node[roundnode] (b0) at (0, 2) {$b^{(t)}$};
		\node[roundnode] (b1) at (5, 2) {$b'$};
		\node[roundnode] (b2) at (10, 2) {$b^{(t+1)}$};
		\node[roundnode] (t0) at (0, 0) {$\theta^{(t)}$};
		\node[roundnode] (t1) at (5, 0) {$\theta'$};
		\node[roundnode] (t2) at (10, 0) {$\theta^{(t+1)}$};

@@

See Appendix \ref{appdx:form-U} for the derivation. The function $U(\theta)$ is a typical objective function for neural network training. The first term is introduced by the likelihood; we collect it into $N \cdot \Lcal(\theta)$, which is the cross-entropy between the graph-predicted and feature-predicted block memberships summed over all vertices. 
The second, introduced by the prior, brings a form of regularisation, guarding against over-fitting. Unlike in many applications where the goal 
is to find the minimiser of $U(\theta)$, our goal is to draw samples from the posterior $\pi_\theta(\cdot) \propto \exp(-U(\cdot))$. We can use $\nabla U$ as a useful heuristic to bias our proposal towards regions of higher target density. We therefore adopt the Metropolis-adjusted Langevin algorithm (MALA) \cite{mala-tweedie}. Given the current sample $\theta$, we generate 
a new sample $\theta'$ as,
%
\begin{equation*}
	\theta' = \theta - h \nabla U(\theta) + \sqrt{2h} \cdot \xi,
\end{equation*}
%
where $\xi \sim \Gaussian(0, I)$ and $h$ is a step-size parameter 
which may vary with the sample index (see Appendix \ref{appdx:step-size} for a discussion of step-size variation).
This leads to the proposal distribution,
\begin{equation*}
	q_\theta(\theta, \theta') \sim \Gaussian \left( \theta' ; \theta - h \nabla U(\theta), 2h I \right).
\end{equation*}
%
Without the injected noise term, MALA is equivalent to gradient descent. We require the noise term $\xi$ to fully explore the parameter space. 
The term $\nabla U$ has an easy to compute analytic form (derived in Appendix \ref{appdx:gradu}). By noting that $\theta = \{w_k\}_{k=1}^{B}$, we write the derivative with respect to each $w_k$ as,
%
\begin{equation}
	\frac{\partial U}{\partial w_k} = - \left( \sum_{i \in [N]} \Big\{ x_i (y_{ik} - a_{ik}) \Big\} - \frac{w_k}{\sigma_\theta^2} \right).
	\label{eqn:U-derivative}
\end{equation}
%
After a proposed move is generated from $q_\theta$,
it is either accepted or rejected in the standard
Metropolis-Hastings accept/reject step.

@@

Up to this point, each $\theta^{(t)}$ update uses its corresponding $b^{(t)}$ sample. This means that the evaluation of $U(\theta)$ and $\nabla U(\theta)$ has high variance. This may lead to longer burn-in for the resulting Markov chain. The only link between $b^{(t)}$ and $\theta^{(t)}$ is in the evaluation of $U(\theta)$ and $\nabla U(\theta)$ which depends only on the matrix $y^{(t)}$ with entries $y_{ij}^{(t)} \coloneqq \one\{b_i^{(t)} = j\}$. We would rather deal with the expectation of each $y_{ij}^{(t)}$:
%
\begin{equation}
	\Expect \left[ y_{ij}^{(t)} \right] = \Expect_{b^{(t)}} \left[ \one(b_{i}^{(t)} = j) \right]
	= p(b_i = j | A, X).
\end{equation}
%
We can obtain an unbiased estimate for this quantity using 
the thinned $b$-samples after burn-in; the length of the
burn-in and the details of the thinning are described
in Appendix \ref{appdx:burn-in-thinning}.
Let $\Tcal_b$  denote the retained set of indices 
for the $b$-samples and $\Tcal_\theta$ similarly for the $\theta$-chain. 
The unbiased estimate for $y_{ij}^{(t)}$ using the 
restricted sample set $\Tcal_b$ is denoted $\hat{y}_{ij}$:
%
\begin{equation}
	\hat{y}_{ij} \coloneqq \frac{1}{|\Tcal_b|} \sum_{t \in \Tcal_b} y_{ij}^{(t)} = \frac{1}{|\Tcal_b|} \sum_{t \in \Tcal_b} \one\{b_i^{(t)} = j\}.
	\label{eqn:y-hat}
\end{equation}
%
The same matrix $\hat{y}$ is used in the update step
for each $\theta^{(t)}$.
This way, it is not necessary to run the $b$ and $\theta$ Markov chains 
concurrently. Instead, we run the $b$-chain to completion and use it 
to generate $\hat{y}$. This affords us the flexibility to vary the 
lengths of the $b$ and $\theta$-chains. Furthermore, the changeover 
from $y^{(t)}$ to $\hat{y}$ reduces the burn-in time for 
the $\theta$-chain by reducing the variance in the evaluation 
of $U$ and $\nabla U$. A detailed description of the overall 
algorithms is given in Appendix \ref{appdx:algorithms}.

@@

Once the samples $\left\{ \theta^{(t)} \right\} \sim p(\theta | A, X)$
have been obtained, we can compute the empirical mean and standard deviation of each component of $\theta$. Switching back to matrix notation, we define $\theta = W$, such that $W_{ij}$ is the weight component for block $i$ and feature $j$. We can then define:
%
\begin{equation}
	\hat{\mu}_{ij} \coloneqq \frac{1}{|\Tcal_\theta|} \sum_{t \in \Tcal_\theta} W_{ij}^{(t)} \qquad \textrm{and} \qquad
	\hat{\sigma}_{ij}^2 \coloneqq \frac{1}{|\Tcal_\theta|} \sum_{t \in \Tcal_\theta} \left( W_{ij}^{(t)} - \hat{\mu}_{ij} \right)^2.
\end{equation}
%
A simple heuristic to discard the least important features requires specifying a cutoff $c > 0$ and a multiplier $k > 0$. We define the function $\Fcal_i(j)$ 
as in~(\ref{eqn:fij}) and only keep features with indices $d \in \Dcal'$, where $\Dcal'$ is given in~(\ref{eqn:kept-feature-set}).
%
\begin{align}
	\Fcal_i(j) &\coloneqq (\hat{\mu}_{ij} - k \hat{\sigma}_{ij}, \hat{\mu}_{ij} + k \hat{\sigma}_{ij}) \cap (-c, +c),
	\label{eqn:fij} \\
	\Dcal' &\coloneqq \left\{ j \in [D] : \exists i \in [B] \textrm{ s.t. }  \Fcal_i(j) = \emptyset \right\}
	\label{eqn:kept-feature-set}
\end{align}
%
Intuitively, this means discarding any feature $j$ for which 
$(\hat{\mu}_{ij} - k\hat{\sigma}_{ij}, \hat{\mu}_{ij} + k \hat{\sigma}_{ij})$ overlaps with
$(-c, c)$ for all block indices $i$. If we were to use the Laplace approximation for the posterior $p(W_{ij} | A, X) \approx \Gaussian(W_{ij}; \hat{\mu}_{ij}, \hat{\sigma}_{ij}^2)$, then this is analogous to a hypothesis test on the value of $W_{ij}$ as in~(\ref{eqn:hyp-test-discard});
then $\Dcal'$ comprises all features $j$ for which $H_1$ is accepted at least once for some $i \in [B]$.
%
\begin{equation}
	H_0: |W_{ij}| \leq c \qquad
	H_1: |W_{ij}| > c
	\label{eqn:hyp-test-discard}
\end{equation}
%
The multiplier $k$ in~(\ref{eqn:fij}) determines the degree of significance of the result. However, as the Laplace approximation is not exact, we only treat this dimensionality reduction method as a useful heuristic and not an exact method. Appendix~\ref{appdx:hyp-test} explores this subject at greater depth.

Conversely, we could fix $k=k_0$ and the dimension of our reduced feature set $|\Dcal'|=D'$. We would then like to find the largest value of $c$ such that $|\Dcal'|=D'$ given $k=k_0$. This is summarised in equation \ref{eqn:c-star}.
%
\begin{equation}
	c^* = \argmax \{c>0\; : \;|\Dcal'| = D', k=k_0\}.
	\label{eqn:c-star}
\end{equation}
%
For an algorithmic implementation of this method refer to appendix \ref{appdx:algorithms}. This approach has the advantage that it fixes the number of reduced dimensions.

>>>>

Having completed the definition of the FFBM, we wish to leverage it 
to perform inference. Given a vertex-labelled graph $(A, X)$,
the goal is to sample from the posterior distribution
of $\theta$:
%
\begin{equation}
	\label{eqn:theta-target}
	\theta \sim p(\theta | A, X).
\end{equation}
%
We propose an iterative approach to obtain samples
$\theta^{(t)}$ from $p(\theta|A,X)$. We first draw samples $b^{(t)}$ 
from the block membership posterior distribution,
and then use each $b^{(t)}$ to obtain a corresponding
sample $\theta^{(t)}$:
%
\begin{align}
	b^{(t)} &\sim p ( b | A, X )  \label{eqn:b-samples},\\
	\theta^{(t)} &\sim p(\theta | X, b^{(t)} ). \label{eqn:theta-samples}
\end{align}
%
Both of these sampling steps can be implemented through a
Markov chain Monte-Carlo approach via a Metropolis-Hastings 
algorithm \cite{hastings-alg}, as a two-level Markov chain.

@@

	\begin{tikzpicture}[
		scale=0.8, every node/.style={transform shape},
		roundnode/.style={circle, draw=black, minimum size=12mm},
		squarednode/.style={rectangle, draw=black, minimum size=12mm}
		]
		% nodes
		\node[roundnode] (b0) at (0, 2) {$b^{(t)}$};
		\node[roundnode] (b1) at (5, 2) {$b'$};
		\node[roundnode] (b2) at (10, 2) {$b^{(t+1)}$};
		\node[roundnode] (t0) at (0, 0) {$\theta^{(t)}$};
		\node[roundnode] (t1) at (5, 0) {$\theta'$};
		\node[roundnode] (t2) at (10, 0) {$\theta^{(t+1)}$};

@@

See Appendix \ref{appdx:form-U} for the derivation. The function $U(\theta)$ is a typical objective function for neural network training. The first term is introduced by the likelihood; we collect it into $N \cdot \Lcal(\theta)$, which is the cross-entropy between the graph-predicted and feature-predicted block memberships summed over all vertices. 
The second, introduced by the prior, brings a form of regularisation, guarding against over-fitting. Unlike in many applications where the goal 
is to find the minimiser of $U(\theta)$, our goal is to draw samples from the posterior $\pi_\theta(\cdot) \propto \exp(-U(\cdot))$. We can use $\nabla U$ as a useful heuristic to bias our proposal towards regions of higher target density. We therefore adopt the Metropolis-adjusted Langevin algorithm (MALA) \cite{mala-tweedie}. Given the current sample $\theta$, we generate 
a new sample $\theta'$ as,
%
\begin{equation*}
	\theta' = \theta - h \nabla U(\theta) + \sqrt{2h} \cdot \xi,
\end{equation*}
%
where $\xi \sim \Gaussian(0, I)$ and $h$ is a step-size parameter 
which may vary with the sample index.
This leads to the proposal distribution,
\begin{equation*}
	q_\theta(\theta, \theta') \sim \Gaussian \left( \theta' ; \theta - h \nabla U(\theta), 2h I \right).
\end{equation*}
%
Without the injected noise term, MALA is equivalent to gradient descent. We require the noise term $\xi$ to fully explore the parameter space. 
The term $\nabla U$ has an easy to compute analytic form (derived in Appendix \ref{appdx:gradu}). By noting that $\theta = \{w_k\}_{k=1}^{B}$, we write the derivative with respect to each $w_k$ as,
%
\begin{equation}
	\frac{\partial U}{\partial w_k} = - \left( \sum_{i \in [N]} \Big\{ x_i (y_{ik} - a_{ik}) \Big\} - \frac{w_k}{\sigma_\theta^2} \right).
	\label{eqn:U-derivative}
\end{equation}
%
After a proposed move is generated from $q_\theta$,
it is either accepted or rejected in the standard
Metropolis-Hastings accept/reject step.

@@

Up to this point, each $\theta^{(t)}$ update uses its corresponding $b^{(t)}$ sample. This means that the evaluation of $U(\theta)$ and $\nabla U(\theta)$ has high variance. This may lead to longer burn-in for the resulting Markov chain. The only link between $b^{(t)}$ and $\theta^{(t)}$ is in the evaluation of $U(\theta)$ and $\nabla U(\theta)$ which depends only on the matrix $y^{(t)}$ with entries $y_{ij}^{(t)} \coloneqq \one\{b_i^{(t)} = j\}$. We would rather deal with the expectation of each $y_{ij}^{(t)}$:
%
\begin{equation}
	\Expect \left[ y_{ij}^{(t)} \right] = \Expect_{b^{(t)}} \left[ \one(b_{i}^{(t)} = j) \right]
	= p(b_i = j | A, X).
\end{equation}
%
We can obtain an unbiased estimate for this quantity using 
the thinned $b$-samples after burn-in.
Let $\Tcal_b$  denote the retained set of indices 
for the $b$-samples and $\Tcal_\theta$ similarly for the $\theta$-chain. 
The unbiased estimate for $y_{ij}^{(t)}$ using the 
restricted sample set $\Tcal_b$ is denoted $\hat{y}_{ij}$:
%
\begin{equation}
	\hat{y}_{ij} \coloneqq \frac{1}{|\Tcal_b|} \sum_{t \in \Tcal_b} y_{ij}^{(t)} = \frac{1}{|\Tcal_b|} \sum_{t \in \Tcal_b} \one\{b_i^{(t)} = j\}.
	\label{eqn:y-hat}
\end{equation}
%
The same matrix $\hat{y}$ is used in the update step
for each $\theta^{(t)}$.
This way, it is not necessary to run the $b$ and $\theta$ Markov chains 
concurrently. Instead, we run the $b$-chain to completion and use it 
to generate $\hat{y}$. This affords us the flexibility to vary the 
lengths of the $b$ and $\theta$-chains. Furthermore, the changeover 
from $y^{(t)}$ to $\hat{y}$ reduces the burn-in time for 
the $\theta$-chain by reducing the variance in the evaluation 
of $U$ and $\nabla U$.

@@

Once the samples $\left\{ \theta^{(t)} \right\} \sim p(\theta | A, X)$
have been obtained, we can compute the empirical mean and standard deviation of each component of $\theta$. Switching back to matrix notation, we define $\theta = W$, such that $W_{ij}$ is the weight component for block $i$ and feature $j$. We can then define:
%
\begin{equation}
	\hat{\mu}_{ij} \coloneqq \frac{1}{|\Tcal_\theta|} \sum_{t \in \Tcal_\theta} W_{ij}^{(t)} \qquad \textrm{and} \qquad
	\hat{\sigma}_{ij}^2 \coloneqq \frac{1}{|\Tcal_\theta|} \sum_{t \in \Tcal_\theta} \left( W_{ij}^{(t)} - \hat{\mu}_{ij} \right)^2.
\end{equation}
%
A simple heuristic to discard the least important features requires specifying a cutoff $c > 0$ and a multiplier $k > 0$. We define the function $\Fcal_i(j)$ 
as in~(\ref{eqn:fij}) and only keep features with indices $d \in \Dcal'$, where $\Dcal'$ is given in~(\ref{eqn:kept-feature-set}).
%
\begin{align}
	\Fcal_i(j) &\coloneqq (\hat{\mu}_{ij} - k \hat{\sigma}_{ij}, \hat{\mu}_{ij} + k \hat{\sigma}_{ij}) \cap (-c, +c),
	\label{eqn:fij} \\
	\Dcal' &\coloneqq \left\{ j \in [D] : \exists i \in [B] \textrm{ s.t. }  \Fcal_i(j) = \emptyset \right\}
	\label{eqn:kept-feature-set}
\end{align}
%
Intuitively, this means discarding any feature $j$ for which 
$(\hat{\mu}_{ij} - k\hat{\sigma}_{ij}, \hat{\mu}_{ij} + k \hat{\sigma}_{ij})$ overlaps with
$(-c, c)$ for all block indices $i$. If we were to use the Laplace approximation for the posterior $p(W_{ij} | A, X) \approx \Gaussian(W_{ij}; \hat{\mu}_{ij}, \hat{\sigma}_{ij}^2)$, then this is analogous to a hypothesis test on the value of $W_{ij}$ as in~(\ref{eqn:hyp-test-discard});
then $\Dcal'$ comprises all features $j$ for which $H_1$ is accepted at least once for some $i \in [B]$.
%
\begin{equation}
	H_0: |W_{ij}| \leq c \qquad
	H_1: |W_{ij}| > c
	\label{eqn:hyp-test-discard}
\end{equation}
%
The multiplier $k$ in~(\ref{eqn:fij}) determines the degree of significance of the result. However, as the Laplace approximation is not exact, we only treat this dimensionality reduction method as a useful heuristic and not an exact method.

Conversely, we could fix $k=k_0$ and the dimension of our reduced feature set $|\Dcal'|=D'$. We would then like to find the largest value of $c$ such that $|\Dcal'|=D'$ given $k=k_0$. This is summarised in equation \ref{eqn:c-star}.
%
\begin{equation}
	c^* = \argmax \{c>0\; : \;|\Dcal'| = D', k=k_0\}.
	\label{eqn:c-star}
\end{equation}
%
This approach has the advantage that it fixes the number of reduced dimensions.
